Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-11 Manuscript/manuscript.tex
Results and discussion updated

\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
  \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
  \caption{\label{simVSexpNOions}
    Top: Chemical structure of POPC with definitions of different order parameters calculated.
    Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  
    from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
    compared with experimental order parameters from \cite{ferreira13}.
    Error bars are approximately the size of the points for head group order parameters.
    Point size for the order parameters of tails are decreased by a factor of 3 to improve the clarity of the plot.
    Bottom: X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
    } 
\end{figure}

@@

The x-ray scattering form factor (Fig.~\ref{simVSexpNOions}) and
the area per lipid (Table~\ref{tab:apls}) of POPC bilayer simulated
with the ECC-lipid model are in good agreement with the experimental results,
as well as with the Lipid14 model. Thus, we conclude that the lipid bilayer dimensions
are well captured by the developed ECC-lipid model. 
Areas per lipid of ECC-lipids with various other water models are documented in SI, table~\ref{tab:apls_si}. 
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.

As in the original Lipid14 model~\cite{dickson14}, the acyl chain order parameters of the
ECC-lipid model agree well with the experimental values in Fig.~\ref{simVSexpNOions}. Notably, the experimentally measured forking and
small order parameter values of $C_2$ segment in {\it sn}-2 chain are relatively well
reproduced by the both models. This has been suggested to indicate that the carbonyl
region of {\it sn}-2 chain is directed towards the water phase, in contrast to the
carbonyl in {\it sn}-1 chain, which would orient more along the bilayer
plane~\cite{seelig75,schindler75,gawrisch92}. While this may be an important
feature for the ion binding details, it is not necessarily reproduced by the
available lipid models~\cite{ollila16}
%This is expected as the ECC model does not modify
%the already highly optimized parameters of this region in Lipid14.

The headgroup order parameters of $\alpha$ and $\beta$ carbons are slightly larger
in the ECC-lipid model than in Lipid14, which is apparently related to the
smaller P-N vector angle with respect to the membrane normal in
Fig.~\ref{OrderParameterCHANGESsurf}. With the current data we cannot,
however, conclude which one of the models give the more realistic
headgroup conformations. The ECC-lipid model gives
the $\beta$ carbon order parameter value closer to experiments, while
value for $\alpha$ carbon is better in Lipid14.
Despite of some deviations from the experimental order parameter values
in Fig. \ref{simVSexpNOions},
the accuracy of the both models in the glycerol backbone region
is comparable to the other state
of art lipid models available in literature \cite{botan15}.

@@

\subsection{Response of POPC head groups to bound charge}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.pdf}
  \caption{\label{OrderParameterCHANGESsurf}
    Headgroup order parameter changes and P-N vector orientation as a function of
    cationic surfactant (dihexadecyldimethylammonium bromide, 2C$_{16}$$^+$N2C$_1$Br$^-$)
    in PC bilayer from simulations and experiments \cite{scherer89}.
  }
\end{figure}

Before proceeding to the ion binding affinity, we quantify
the response of headgroup order parameters to the amount of 
bound charge by using mixtures of monovalent cationic surfactants
and POPC~\cite{scherer89}. The amount of bound charge per PC 
in these systems is given by the molar fraction of cationic 
surfactants, because essentially all surfactants locate in the 
lipid bilayers. Experimental data for these systems can be used to validate 
the sensitivity of lipid headgroup order parameters
to the amount of bound charge in simulations.

%depends on both, the
%binding affinity and the response of the head group on bound charge.
%Thus, the head group response to bound charge has to be correctly
%described in the model if the concept is used for quantitative studies of
%binding affinity. Hence, we first quantify the response of head group order
%parameters to the bound charge by using experimental data measured from
%mixtures of cationinc surfactants and PC lipids~\cite{scherer89}.
%
%Thus, the head group order parameter response to bound charge can be first
%quantified aginst experimental data before proceeding to binding affinity studies.
%Headgroup order parameter response to the amount of bound charge can
%be quantified by using ionic surfactants with a known valency~\cite{scherer89}.
%All surfactants locate in lipid bilayers due to their amphiphilic nature,
%thus the exact amount of bound charge in the membrane is known in these systems.
%The head group order parameter changes as a function of bound charges can be then
%explicitly determined from such systems.

The headgroup order parameter changes with an increasing amount of 
the cationic surfactant dihexadecyldimethylammonium bromide 
is compared between experiments~\cite{scherer89} and simulations in
Fig. \ref{OrderParameterCHANGESsurf}. 
The observed order parameter decrease in simulations and experiments
can be approximated to be linear at least for mole fractions
below $\sim$30\%, as expected from Eq. \ref{OPchangeEQ}.
The slope is, however, too steep in the Lipid14 model indicating that 
the head group order parameters are too sensitive to a bound charge.
The ECC-lipids model gives a slope in very good agreement with experiments
for the $\alpha$ segment, while the slope is slightly
underestimated for the $\beta$ segment.
%This has to be taken into account when interpreting the
%head group order parameter response to CaCl$_2$ concentration in Lipid14 model.
%observed in  Ref. \citenum{catte16}, see also Fig. \ref{fig:delta_ordPar_NaCl}.
%The present comparison of head group order parameter changes 
%suggests that the observed overestimated response of all models to 
%increasing CaCl$_2$ concentration in \cite{catte16} 
%can be explained at least in part by a high sensitivity of the head group order parameter response
%to the bound charge.
\todo{SAMULI: We could calculate the slopes from simulations, but I am not sure if
we would actually learn anything useful from this.}

The headgroup P-N vector angle with respect to the membrane normal
is also shown as a function of cationic surfactant mole fraction
in Fig.~\ref{OrderParameterCHANGESsurf}.
The headgroup orients more towards the water phase with an increasing
amount of bound cations, as previously reported in Ref.~\citenum{seelig87}. 
The effect is more pronounced in Lipid14 than 
in ECC-lipids model, which is in line with the order parameter 
results and the reduced charge--dipole interactions in the latter model.
The response of $\alpha$-order parameter to bound positive charge
in ECC-lipid model is in good agreement with experiments.
The model can be thus used to study changes of lipid P-N vector
in varying conditions.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
  \caption{\label{fig:delta_ordPar_CaCl}
    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
    are shown from simulations with different force fields together with experimental data 
    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
  }
\end{figure}

The binding affinities of aqueous cations to lipid 
bilayers can be measured using the headgroup 
order parameters, because they decrease proportionally  
to the bound positive charge~\cite{seelig87,catte16}. 
The headgroup order parameter responses 
to increasing aqueous CaCl$_2$ concentrations 
from experiments (DPPC \cite{akutsu81} 
and POPC \cite{altenbach84}) and different simulation models for POPC 
are shown in Fig.~\ref{fig:delta_ordPar_CaCl}.
Responses to NaCl concentrations are in SI, Fig.~S\ref{fig:delta_ordPar_NaCl}.

%The simulation results of Lipid14 with \AA{}qvist ions~\cite{aqvist90} (from Ref. \citenum{catte16})
%are shown together with the results from Lipid14 simulation
%with ions by Dang~\cite{smith94,chang1999,dang2006} and from ECC-lipids simulations with ECC-ions.

Negligible changes of the headgroup order parameters are measured
with submolar concentrations of NaCl due to the very low affinity of Na$^+$
in PC bilayers \cite{akutsu81}. While Na$^+$ binding and the related 
headgroup order parameter changes were overestimated in
almost all the available simulation models, the low affinity
and negligible order parameter changes were reproduced by Lipid14 
model when simulated with \AA{}qvist ions~\cite{catte16}.
However, the same combination of force field parameters
overestimated the headgroup order parameter response to CaCl$_2$ 
concentration, which was the case also in all other models tested in
Ref.~\citenum{catte16}.
Using the ion model by Dang et al.~\cite{smith94,chang1999,dang2006} or
ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
with more realistic bulk behaviour did not improve the results
for the CaCl$_2$ interactions with a POPC bilayer described by the Lipid14 model, as seen 
in Figs. \ref{fig:delta_ordPar_CaCl} and S\ref{fig:ordPars_waterModels} (in SI), respectively.
\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
The results support the conclusion of the previous work \cite{catte16}
that improvements also in lipid models are needed to 
correctly describe divalent cation binding to PC bilayers.

Significant improvement can be achieved using the ECC
approach also for lipids. The headgroup order parameters change
as a function of CaCl$_2$ concentration for the ECC-lipid model with ECC-ions
exhibit a good agreement with experiments in Fig.~\ref{fig:delta_ordPar_CaCl} and S\ref{fig:delta_ordPar_NaCl}.
As discussed in previous section, the model gives also
a good agreement with experiments for
the headgroup response to bound charge.
%Thus, the model can used for more detailed analysis of the binding affinity.

The binding affinities are quantified using
the water density profiles along the membrane normal shown
in Fig.~\ref{fig:cacl-dens}. The density profiles show
a larger Ca$^{2+}$ density peak in lipid headgroup region for
the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
The relative surface excess calculated from Eq.~\ref{surfexcess} is 
$\Gamma_i^w = 0.07 \pm 0.01 nm^{-2}$ for the ECC-lipid model,
which is significantly smaller than $\Gamma_i^w = 0.13 \pm 0.01 nm^{-2}$ for
Lipid14 with \AA{}qvist and $\Gamma_i^w = 0.3 \pm 0.03 nm^{-2}$ with Dang ions.

\todo{Below analysis is done in a stupid way to get some idea. I would fine it useful to
  do this analysis by using the density profiles, but it is not necessary.
  The reason why I want to do something like this, is that it would be useful to have
  some easier way to estimate the correct binding affinity than the electrometer concept,
  which is quite tedious to apply in practise (simulations with different concenrations,
  cationic surfactant check, etc.). I wuold like to be able to estimate much faster from a
  simulation if the binding affinity is reasonable or not. This may not be the correct way for
that anyway.}
Rough estimates for the free energy difference between bound and unbound cation
are given by
\begin{equation}
  \Delta G = -{\rm k_b T}\log(\frac{p_i}{p_o}),
\end{equation}
where $p_o$ and $p_i$ are estimated from the Ca$^{2+}$densities in bulk water and
in the maximum density in bilayer, respectively. The density profiles in
Fig. \ref{fig:cacl-dens} give $\sim 0.8$ k$_{\rm b}$T for the free energy difference
between bound and unbound Ca$^{2+}$ ions in ECC-lipid model and $\sim 1.4$ k$_{\rm b}$T
in Lipid14 with \AA{}qvist.

@@

Since the lipid headgroup order parameter responds to the amount of bound charge
and to the aqueous ion concentrations are both in good agreement with experiments
in the ECC-lipid model with ECC-ion parameters, we consider the Na$^+$ and Ca$^{2+}$
binding affinities to be realisticlly described in this model. In contrast, the Ca$^{2+}$
binding affinity is overestimated by Lipid14 model when simulated with all the
tested ion models. Similar conclusions were previously made based only on the headrogup
order parameter data with aqueous cations~\cite{catte16}. However,
the discrepancies with experiments in previous work could partly arise also from
the inaccurately described sensitivity of the headgroup to bound charge. Here we quantify this
effect and conclude that the improvement due to ECC is to a large extent, but not entirely,
caused by a more realistic headgroup sensitivity to bound charge.
This indicates that the issue should be carefully considered also when
the electrometer concept is used to compare ion binding between experiments
and simulations with other models.

%The good agreement with experiments indicates
%that the ECC corrected lipid model has a sufficient accuracy for realistic
%studies of details of ions binding to lipid bilayers.

>>>>

\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
  \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
  \caption{\label{simVSexpNOions}
    Top: The chemical structure of POPC and the labeling of the carbon segments.
    Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipid models
    compared with the experimental values from \cite{ferreira13}.
    The size of the points for the head group order parameters correspond the
    of the error bars ($\pm 0.02$ for experiments \cite{botan15,ollila16} and ?? for simulations).
    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
    Bottom: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
    the ECC-lipid models compared with experiments~\cite{Kucerka2011}.
  } 
  \todo{Increase the height of the order parameter figure} \\
  \todo{sn-1 and sn-2 order parameters should be somehow labelled in the order parameter figure. Maybe empty and filled points?} \\
  \todo{Add size of the error bars in simulations to the caption.} \\
  \todo{Would it be possible to increase the size of points for acyl chain order parameters only in z-direction such that
    it would correspond the error bars?} \\
  \todo{x-axis in form factor plot from 0 to 0.6, where experimental data ends. y-axis from 0 to $\sim$2.5 to remove the empty space.} \\
\end{figure}

@@

The ECC-lipid and Lipid14 models reproduce the experimental x-ray scattering form factors
of POPC bilayer with the similar accuracy in Fig.~\ref{simVSexpNOions}.
The area per lipid values from the Lipid14 model is slightly larger than the
experimental value in Table~\ref{tab:apls}, while the value from ECC-lipid model
is slightly smaller. Area per lipid values of the ECC-lipid model show some variation
when simulated with different water models in Table~\ref{tab:apls_si} (Supplementary Information),
however, all the values are close to the experimentally reported values.
In conclusion, the ECC-lipid model reproduces the experimental dimensions of POPC
lipid bilayer with the accuracy comparable to the other state of the art lipid models \cite{ollila16}.
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.

The acyl chain order parameters of the Lipid14 model~\cite{dickson14} and the
ECC-lipid model agree with the experimental values with in the error bars
in Fig.~\ref{simVSexpNOions}, althought the ECC-lipid model gives slightly larger
values for {\it sn}-1 chain. Notably, the experimentally measured forking and
small order parameter values of $C_2$ segment in {\it sn}-2 chain are relatively well
reproduced by the both models. This has been suggested to indicate that the carbonyl
of {\it sn}-2 chain is directed towards the water phase, in contrast to the
carbonyl in {\it sn}-1 chain, which would orient more along the bilayer
plane~\cite{seelig75,schindler75,gawrisch92}. While this may be an important
feature for the ion binding details, it is not necessarily reproduced by the
available lipid models~\cite{ollila16}
%This is expected as the ECC model does not modify
%the already highly optimized parameters of this region in Lipid14.

The order parameters of $\alpha$ and $\beta$ carbons in the headgroup are slightly larger
in the ECC-lipid model than in the Lipid14 model, which is apparently related to the
P-N vector orienting 7$^{\circ}$ more toward the water phase in the ECC-lipid model, 
see Fig.~\ref{OrderParameterCHANGESsurf}. With the current data we cannot,
however, conclude which one of the models give the more realistic
headgroup conformations. The ECC-lipid model gives
the $\beta$ carbon order parameter value closer to experiments, while
value for $\alpha$ carbon is better in the Lipid14 model.
Despite of some deviations from the experimental order parameter values
in Fig. \ref{simVSexpNOions},
the accuracy of the both models in the glycerol backbone region
is comparable to the other state
of art lipid models available in literature \cite{botan15}.

@@

\subsection{Response of POPC head groups to bound charge}\label{section:boundCHARGE}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.pdf}
  \caption{\label{OrderParameterCHANGESsurf}
    The changes of headgroup order parameters and P-N vector orientation as a function of
    cationic surfactant (dihexadecyldimethylammonium) in POPC bilayer from simulations
    and experiments \cite{scherer89}.
  }
  \todo{Labeling should be consistent with the previous figure, i.e. experimental data with diamonds.} \\
  \todo{x-axis scale from -1 to 51 to make the point in zero fully visible and to remove empty space.} \\
  \todo{Empty space between figures and from the right column could be reduced.}
\end{figure}

Before proceeding to the ion binding affinity, we quantify
the response of the headgroup order parameters to the amount of 
bound charge by using mixtures of monovalent cationic
surfactants (dihexadecyldimethylammonium)
and POPC~\cite{scherer89}. The amount of bound charge per PC 
in these systems is given by the molar fraction of cationic 
surfactants, because essentially all surfactants with two hydrophopic
acyl chains can be assumed to locate in the lipid bilayers.
The experimental data for these systems can be used to validate 
the sensitivity of lipid headgroup order parameters
to the amount of bound charge in simulations.

The changes of the headgroup order parameters with an increasing amount of 
the cationic surfactant is compared between experiments~\cite{scherer89} and
simulations in Fig. \ref{OrderParameterCHANGESsurf}.
Approximately linear decrease of the order parameters, as expected from Eq. \ref{OPchangeEQ},
is observed in simulations and experiments at least for the mole fractions
below $\sim$30\%. The slope is, however, too steep in the Lipid14 model indicating that 
the head group order parameters respond too sensitively to the bound positive charge.
The slope in the ECC-lipid model is in very good agreement with experiments
for the $\alpha$ segment, while the slope is slightly
underestimated for the $\beta$ segment.

The headgroup P-N vector angle with respect to the membrane normal
is also shown in Fig.~\ref{OrderParameterCHANGESsurf}
as a function of the mole fraction of the cationic surfactant.
As suggested previously~\cite{seelig87}, the headgroup orients
more towards the water phase with the increasing amount of positive
charge in a PC lipid bilayer. The effect is more pronounced in the
Lipid14 model, for which the addition of 50\% mole fraction of the cationic
surfactant leads to the decrease of 20$^{\circ}$ of the P-N vector angle,
while the corresponding change in the ECC-lipid model is 11$^{\circ}$.
The difference is in line with the smaller order parameter 
changes and the reduced charge--dipole interactions in the latter model.
The lesser sensitivity of the P-N vector angle response in the ECC-lipid
model can be considered to be more realistic, because the changes of the headgroup
order parameters as a function of the bound positive charge
are in better agreement with experiments in this model.
The results also imply that the validation and improvements of MD simulation
models are generally required for the reliable simulation studies of the
lipid headgroup responses to ions or other biomolecules, as also concluded
previously \cite{botan15}.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
  \caption{\label{fig:delta_ordPar_CaCl}
    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
    are shown from simulations with different force fields together with experimental data 
    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
  }
  \todo{I think that we must show the NaCl results in the main paper. I think that the best would
    be to use similar two column format, which we had before. This is used also in NMRlipids II and makes the
    comparison easier} \\
  \todo{The DPPC and POPC should be shown and properly labelled as we had previously.} \\
  \todo{The labeling should be consistent with previous figures, i.e., experiments with diamonds.} \\
  \todo{Start x-axis scale from -1 to make the point in zero fully visible.} \\
  \todo{Start y-axis scale of the two top figures from -0.13 to remove the empty space.} \\
  \todo{y-axis scale of the bottom figure from 59 to 78 to remove the empty space.} \\
  \todo{Empty space between figures and from the right column could be reduced.} \\
  \todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
    the value of $g_3$ order parameter of DPPC was -0.214 
    in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
    of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
    at the C-2 segments of both fatty acyl chains.''
  }
\end{figure}

The changes of the lipid bilayer headgroup order parameter from different simulations and
experiments \cite{akutsu81,altenbach84} are shown in Fig.~\ref{fig:delta_ordPar_CaCl}
as a function of NaCl and CaCl$_2$ concentrations.
%Responses to NaCl concentrations are in SI, Fig.~S\ref{fig:delta_ordPar_NaCl}.
These results can be used to compare the ion binding affinities to lipid bilayers
between simulations and experiments by using the electrometer concept, because 
the order parameters decrease proportionally to the amount of bound positive charge~\cite{seelig87,catte16}. 
The recent comparison of different simulation models to the experimental data
revealed that most models significantly overestimate the Na$^+$ ion binding to PC
lipid bilayers and that none of the available models correctly reproduce
the details of the binding of Ca$^{2+}$ ions \cite{catte16}.
A positive exception was the Lipid14 model \cite{dickson14} simulated with \AA{}qvist ions,
as seen from the data replotted from Ref.~\citenum{catte16} in Fig.~\ref{fig:delta_ordPar_CaCl}.
The model reproduced the experimentally measured small order parameter changes with NaCl
and imperceptible Na$^+$ binding to PC bilayers \cite{akutsu81,altenbach84}.
However, the changes of the headgroup order parameters as a function of CaCl$_2$ 
concentration were overestimated by the same combination of force field parameters.

Since significant artefacts are reported in simulations with \AA{}qvist ions is water \cite{??},
we also simulated the Lipid14 model with the ion models by Dang et al.~\cite{smith94,chang1999,dang2006} and
ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014} having
more realistic bulk behaviour. Instead of the improvement in the binding behaviour,
we observed overbinding also of Na$^+$ with these ions models, while the results
with Ca$^{2+}$ were barely affected,
%for the CaCl$_2$ interactions with a POPC bilayer described by the Lipid14 model,
as seen in Figs. \ref{fig:delta_ordPar_CaCl} and S??.%\ref{fig:ordPars_waterModels} (in SI), respectively.
\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
The results are in line with the previous work \cite{catte16},
suggesting that the improvements in the lipid parameters are
required to correctly describe the divalent cation binding to PC lipid bilayers.

To improve the cation binding behaviour to POPC lipid bilayer we implicitly included the electronic
polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
The ECC-lipid model simulations with the ECC-ion models \cite{??}
exhibit a good agreement with the experimental changes of 
the lipid headgroup order parameters as a function of NaCl and CaCl$_2$ concentrations
in Fig.~\ref{fig:delta_ordPar_CaCl}. Since also the headgroup order parameter
response the to the bound positive charge the ECC-lipid model
was in good agreement with experiments in section \ref{section:boundCHARGE},
we conclude that the model correctly reproduces the binding affinity of
Na$^{+}$ and Ca$^{2+}$ ions to POPC lipid bilayer.

The density distributions of water and ions along the membrane normal
are shown in Fig.~\ref{fig:cacl-dens}.
\todo{Densities and relative surface excess from NaCl systems to be added.
  The discussion below to be finished after this.}
%The density profiles show
%a larger Ca$^{2+}$ density peak in lipid headgroup region for
%the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
To quantify the binding affinities in different models we calculated 
the relative surface excess from Eq.~\ref{surfexcess}. The value for Calcium
in the ECC-lipid model, $\Gamma_i^w = 0.07 \pm 0.01 \rm{nm}^{-2}$, is significantly
smaller than the values for the Lipid14 model with \AA{}qvist ions, $\Gamma_i^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
or with Dang ions, $\Gamma_i^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.

%\todo{Below analysis is done in a stupid way to get some idea. I would fine it useful to
%  do this analysis by using the density profiles, but it is not necessary.
%  The reason why I want to do something like this, is that it would be useful to have
%  some easier way to estimate the correct binding affinity than the electrometer concept,
%  which is quite tedious to apply in practise (simulations with different concenrations,
%  cationic surfactant check, etc.). I wuold like to be able to estimate much faster from a
%  simulation if the binding affinity is reasonable or not. This may not be the correct way for
%that anyway.}
%Rough estimates for the free energy difference between bound and unbound cation
%are given by
%\begin{equation}
%  \Delta G = -{\rm k_b T}\log(\frac{p_i}{p_o}),
%\end{equation}
%where $p_o$ and $p_i$ are estimated from the Ca$^{2+}$densities in bulk water and
%in the maximum density in bilayer, respectively. The density profiles in
%Fig. \ref{fig:cacl-dens} give $\sim 0.8$ k$_{\rm b}$T for the free energy difference
%between bound and unbound Ca$^{2+}$ ions in ECC-lipid model and $\sim 1.4$ k$_{\rm b}$T
%in Lipid14 with \AA{}qvist.

In conclusion, the results suggest that 
the ECC-lipid model together with the ECC-ion parameters gives a realistic
description for the binding affinities of Na$^+$ and Ca$^{2+}$ ions to POPC lipid
bilayer. Furthermore, the overestimated headgroup order parameter changes of POPC in the Lipid14
model as a function of CaCl$_2$ concentration arise from both, the overestimated
binding affinity and sensitivity of the headgroup tilt to the bound positive charge.
This probably applies also to the other lipid models considered in the previous study~\cite{catte16},
emphasizing the importance of the comparison of the lipid headgroup order parameter response
to the bound charge between simulations and experiments as done in section \ref{section:boundCHARGE}.

@@
